---
title: "scRNA-glioma-workflows: van Hijfte dataset - Sample Y"
author: 
  - Youri Hoogstrate
  - Levi van Hijfte
  - Santoesha A. Ghisai
  - Maud Diepeveen
date: "`r BiocStyle::doc_date()`"
---

# Load libraries and functions

```{r}
library(ggplot2)

#source('R/function__remove_doublets.R')
source('R/function__annotate_read_fractions.R')
source('R/function__cache_seurat_object.R')
source('R/function__celltype_reclassification_PSLDA.R')
source('R/function__flag_uncertain_classifications.R')

source('R/load__celltype_markers.R')
```


# Import data & set config

```{r}

config <- list(
  sample_name = 'SF11136',
  sample_full_name = 'GSM4119535_SF11136_LGG',
  path = "data/GSE138794_Diaz/scRNA_GSM4119535_SF11136_LGG/",
  
  random.seed = 42, # don't ask me but you can guess why, its Seurats default
  
  seurat.min.cells = 3,
  seurat.min.features = 200,
  
  seurat.vst.nfeatures = 2000,
  
  min.nFeature_RNA = 750,
  min.nCount_RNA = 1400,
  
  pca.informative.components = 21,
  
  cluster.resolution = 1.2
)


obj__diaz_2019__SF11136 <- Seurat::Read10X(
  data.dir = config$path,
  gene.column=1
) |> Seurat::CreateSeuratObject(
  min.cells = config$seurat.min.cells,
  min.features = config$seurat.min.features,
  project = config$sample_name
)

obj__diaz_2019__SF11136@misc <- config
obj__diaz_2019__SF11136$celltype.annotated <- NA



obj__diaz_2019__SF11136@misc$reclass.PSLDA.min.scores <- list(
  'oligodendrocyte' = 0.365,
  'tumor' = 0.33,
  'neuron' = 0,
  'endothelial' = 0,
  'pericyte' = 0.22,
  'T-cell' = 0
)


rm(config)

set.seed(obj__diaz_2019__SF11136@misc$random.seed)
stopifnot(.Random.seed[1] == 10403) # I guess this is some universal thing?
```

# Annotate fractions rRNA and MT-RNA

```{r}
obj__diaz_2019__SF11136 <- annotate_read_fractions(obj__diaz_2019__SF11136) # percentage MT, Ribo RNA and MALAT1



obj__diaz_2019__SF11136$percentage_mitochondrial_reads_scale <- log((obj__diaz_2019__SF11136$percentage_mitochondrial_reads * 500) + 1)

obj__diaz_2019__SF11136$percentage_MALAT1_reads_scale <- log((obj__diaz_2019__SF11136$percentage_mitochondrial_reads * 500) + 1)

obj__diaz_2019__SF11136$percentage_ribosomal_reads_scale <- log((obj__diaz_2019__SF11136$percentage_ribosomal_reads * 100) + 1)


obj__diaz_2019__SF11136[['nCount_RNA_scale']] <- log(obj__diaz_2019__SF11136$nCount_RNA / 100)

obj__diaz_2019__SF11136[['nFeature_RNA_scale']] <- log(obj__diaz_2019__SF11136$nFeature_RNA / 100)



```

# Q/C
## Exclude low-res cells

```{r}
Seurat::VlnPlot(object = obj__diaz_2019__SF11136, features = c("percentage_ribosomal_reads", "percentage_MALAT1_reads"), ncol = 3, pt.size = 0.01, group.by = "orig.ident")

Seurat::VlnPlot(object = obj__diaz_2019__SF11136, features = c("nFeature_RNA", "nCount_RNA","percentage_mitochondrial_reads"), ncol = 2, pt.size = 0.01, group.by = "orig.ident")
```


```{r}
ggplot(obj__diaz_2019__SF11136@meta.data, aes(y=`nFeature_RNA`, x=orig.ident)) +
  geom_jitter(cex=0.01) +
  geom_hline(yintercept = obj__diaz_2019__SF11136@misc$min.nFeature_RNA, col="red") +
  geom_hline(yintercept = 4500,col="red")

ggplot(obj__diaz_2019__SF11136@meta.data, aes(y=`nCount_RNA`, x=orig.ident)) +
  geom_jitter(cex=0.01)  +
  geom_hline(yintercept = obj__diaz_2019__SF11136@misc$min.nCount_RNA, col="red") +
  geom_hline(yintercept = 14000,col="red")
  #scale_y_log10()

obj__diaz_2019__SF11136 <- subset(x = obj__diaz_2019__SF11136, subset = 
                     nFeature_RNA >= obj__diaz_2019__SF11136@misc$min.nFeature_RNA &
                     nCount_RNA >= obj__diaz_2019__SF11136@misc$min.nCount_RNA
                     #nFeature_RNA < 4500 & 
                     #nCount_RNA < 14000 &
                     #percent.mito < 0.025
)
```


# Normalize and cluster/UMAP
## Scale / normalize

```{r}
set.seed(obj__diaz_2019__SF11136@misc$random.seed)
obj__diaz_2019__SF11136 <- Seurat::NormalizeData(
  object = obj__diaz_2019__SF11136, 
  normalization.method = "LogNormalize", 
  scale.factor = 1e4,
  verbose = F
  )

set.seed(obj__diaz_2019__SF11136@misc$random.seed)
obj__diaz_2019__SF11136 <- Seurat::FindVariableFeatures(
  object = obj__diaz_2019__SF11136,
  selection.method = "vst",
  nfeatures = obj__diaz_2019__SF11136@misc$seurat.vst.nfeatures,
  verbose = F
)


# Don't do this -- every update of the marker genes file will trigger new umap (+ annotation)
# Append known marker genes to the variable features
# append <- intersect(setdiff(celltype_markers$markers, obj__diaz_2019__SF11136@assays$RNA@var.features), rownames(obj__diaz_2019__SF11136))
# if(length(append) > 0) {
#   message(paste0("Appending ", length(append), " marker genes to variable features"))
#   obj__diaz_2019__SF11136@assays$RNA@var.features <- c(obj__diaz_2019__SF11136@assays$RNA@var.features, append)
# }
# rm(append)


obj__diaz_2019__SF11136 <- Seurat::ScaleData(
  object = obj__diaz_2019__SF11136,
  verbose = T
  )
```

## PCA

```{r}

set.seed(obj__diaz_2019__SF11136@misc$random.seed)
obj__diaz_2019__SF11136 <- Seurat::RunPCA(
  reduction.key = "PC_",
  object = obj__diaz_2019__SF11136,
  features = Seurat::VariableFeatures(object = obj__diaz_2019__SF11136),
  verbose = F,
  seed.use = obj__diaz_2019__SF11136@misc$random.seed
)


Seurat::ElbowPlot(
  object = obj__diaz_2019__SF11136,
  ndims = round(obj__diaz_2019__SF11136@misc$pca.informative.components * 4/3) + 1
)


# @todo add marker colors
Seurat::VizDimLoadings(
  object = obj__diaz_2019__SF11136,
  dims = 1:6,
  reduction = "pca"
)


Seurat::DimPlot(obj__diaz_2019__SF11136, reduction = "pca")



# add pca with reversely signed loadings and reversely signated values
obj__diaz_2019__SF11136@reductions$invpca <- obj__diaz_2019__SF11136@reductions$pca
obj__diaz_2019__SF11136@reductions$invpca@feature.loadings <- obj__diaz_2019__SF11136@reductions$invpca@feature.loadings * -1
obj__diaz_2019__SF11136@reductions$invpca@cell.embeddings <- obj__diaz_2019__SF11136@reductions$invpca@cell.embeddings * -1
Seurat::Key(obj__diaz_2019__SF11136@reductions$invpca) <- "PCinv_"

```


## Clustering

```{r}

set.seed(obj__diaz_2019__SF11136@misc$random.seed)
obj__diaz_2019__SF11136 <- Seurat::FindNeighbors(
  object = obj__diaz_2019__SF11136,
  dims = 1:obj__diaz_2019__SF11136@misc$pca.informative.components,
  verbose = F
)



set.seed(obj__diaz_2019__SF11136@misc$random.seed)
obj__diaz_2019__SF11136 <- Seurat::FindClusters(
  object = obj__diaz_2019__SF11136,
  resolution = obj__diaz_2019__SF11136@misc$cluster.resolution,
  algorithm = 1,
  random.seed = obj__diaz_2019__SF11136@misc$random.seed,
  verbose = F
)


Seurat::DimPlot(obj__diaz_2019__SF11136, reduction = "pca", label = TRUE, pt.size = .6, group.by = "seurat_clusters")




```


## UMAP

```{r}

set.seed(obj__diaz_2019__SF11136@misc$random.seed)
obj__diaz_2019__SF11136 <- Seurat::RunUMAP(
  object = obj__diaz_2019__SF11136,
  dims = 1:obj__diaz_2019__SF11136@misc$pca.informative.components,
  seed.use = obj__diaz_2019__SF11136@misc$random.seed,
  verbose = T
)


Seurat::DimPlot(obj__diaz_2019__SF11136, reduction = "umap")


```

# Annotate cells manually

## TAM

```{r}
Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = celltype_markers |>
    dplyr::filter(celltype == "TAM") |> 
    dplyr::pull(markers) |> 
    head(n = 4),
  label=T
)
```


```{r}
obj__diaz_2019__SF11136$celltype.annotated = ifelse(obj__diaz_2019__SF11136$seurat_clusters %in% c(10),"TAM",obj__diaz_2019__SF11136$celltype.annotated)


Seurat::DimPlot(obj__diaz_2019__SF11136, reduction = "umap", label = TRUE, pt.size = .6, group.by = "celltype.annotated")

```


## oligodendrocyte

```{r}
Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = celltype_markers |>
    dplyr::filter(celltype == "oligodendrocyte") |> 
    dplyr::pull(markers) |> 
    head(n = 4),
  label=T
)


obj__diaz_2019__SF11136$celltype.annotated = ifelse(obj__diaz_2019__SF11136$seurat_clusters %in% c(7,1,2,0,6,9,12),"oligodendrocyte",obj__diaz_2019__SF11136$celltype.annotated)


Seurat::DimPlot(obj__diaz_2019__SF11136, reduction = "umap", label = TRUE, pt.size = .6, group.by = "celltype.annotated")

```


## tumor

```{r}

Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = c('CD44'),
  label=T
)


Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = celltype_markers |>
    dplyr::filter(celltype == "tumor") |> 
    dplyr::pull(markers) |> 
    head(n = 4),
  label=T
)
```


```{r}
obj__diaz_2019__SF11136$celltype.annotated = ifelse(obj__diaz_2019__SF11136$seurat_clusters %in% 
                                            c(4, 8, 5,  3),
                                          "tumor",obj__diaz_2019__SF11136$celltype.annotated)


Seurat::DimPlot(obj__diaz_2019__SF11136, reduction = "umap", label = TRUE, pt.size = .6, group.by = "celltype.annotated")

```


## OPC

```{r, find}

Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = celltype_markers |>
    dplyr::filter(celltype == "OPC") |> 
    dplyr::pull(markers) |> 
    head(n = 8),
  label=T
)

```


```{r, set and confirm}

```

## astrocyte

```{r}
Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = celltype_markers |>
    dplyr::filter(celltype == "astrocyte") |> 
    dplyr::pull(markers) |> 
    head(n = 4),
  label=T
)
```


```{r}


```


## neuron

```{r}
Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = celltype_markers |>
    dplyr::filter(celltype == "neuron") |> 
    dplyr::pull(markers) |> 
    head(n = 6),
  label=T
)


```



## endothelial & pericytes
PC_6 negative = EN
PC_6 positive = PE

```{r}
Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = celltype_markers |>
    dplyr::filter(celltype == "endothelial") |> 
    dplyr::pull(markers) |> 
    head(n = 10),
  label=T
)
```


```{r}
en_pe = obj__diaz_2019__SF11136[,obj__diaz_2019__SF11136$seurat_clusters %in% c(11)]

plot(density(as.data.frame(en_pe@reductions$pca@cell.embeddings)[["PC_6"]]))
abline(v=12)

# PE
as.data.frame(en_pe@reductions$pca@cell.embeddings) |>
  dplyr::filter(PC_6 > 12) |> 
  tibble::rownames_to_column('umi') |> 
  dplyr::pull(umi)

pe.umis <- c("AAGGCAGAGTCCATAC_1", "ATAGACCCAGCCACCA_1", "CGAGCACTCGCATGAT_1", "CGGGTCAAGGTGCTTT_1", "GACGCGTCACCCAGTG_1",
  "TCAATCTCACGCCAGT_1", "TATCTCAAGATCCCAT_2")



# EN
as.data.frame(en_pe@reductions$pca@cell.embeddings) |>
  dplyr::filter(PC_6 <= 12) |> 
  tibble::rownames_to_column('umi') |> 
  dplyr::pull(umi)

en.umis <- c("ACGAGGATCAGAAATG_1", "AGTGTCAAGAGACGAA_1", "ATAGACCCATCGATTG_1", "CCACCTAGTTCCTCCA_1", "CGTTAGAGTTGAGGTG_1", 
"GCGCCAACATCGGGTC_1", "GCTGCGAGTCGCATCG_1", "GGGATGATCGCCTGAG_1", "GTGAAGGAGTCGATAA_1", "TAAGAGATCCGTAGGC_1",
"TAGCCGGAGATCCTGT_1", "ACCCACTAGCAATCTC_2", "AGGGTGACACCAGATT_2", "CAGGTGCAGAATAGGG_2", "CCATTCGTCTCACATT_2",
"CGGACTGCATTGAGCT_2", "CGGCTAGAGAGGGATA_2", "GCAATCAGTGTAACGG_2", "GCGCCAACATCGGGTC_2", "GCTTGAACATGACGGA_2",
"GGCGACTGTCTCGTTC_2", "TCTTTCCGTAGGAGTC_2")

obj__diaz_2019__SF11136$celltype.annotated = ifelse(colnames(obj__diaz_2019__SF11136) %in% en.umis,"endothelial", obj__diaz_2019__SF11136$celltype.annotated)
rm(en.umis)

obj__diaz_2019__SF11136$celltype.annotated = ifelse(colnames(obj__diaz_2019__SF11136) %in% pe.umis,"pericyte", obj__diaz_2019__SF11136$celltype.annotated)
rm(pe.umis)

rm(en_pe)


Seurat::DimPlot(obj__diaz_2019__SF11136, reduction = "umap", label = TRUE, pt.size = .6, group.by = "celltype.annotated")
```


```{r}
Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = celltype_markers |>
    dplyr::filter(celltype == "pericyte") |> 
    dplyr::pull(markers) |> 
    head(n = 10),
  label=T
)
```



## T-cell

```{r}
Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = celltype_markers |>
    dplyr::filter(celltype == "T-cell") |> 
    dplyr::pull(markers) |> 
    head(n = 4),
  label=T
)
```


## Find remaining

```{r, C3}
m.3 <- Seurat::FindMarkers(obj__diaz_2019__SF11136, ident.1 = 3)
head(m.3, n=25)
```
LINC01088  0.000000e+00  2.8359079 0.956 0.026  0.000000e+00
ID4        0.000000e+00  3.0104386 0.973 0.040  0.000000e+00
EDNRB      0.000000e+00  2.5992103 0.938 0.025  0.000000e+00
SLC14A1    0.000000e+00  2.4181379 0.867 0.010  0.000000e+00
GJA1      1.429890e-300  3.5274009 0.894 0.031 2.140688e-296
SLC4A4    3.452693e-297  1.5680509 0.788 0.009 5.169027e-293
MGST1     1.176369e-294  1.7258641 0.779 0.008 1.761143e-290
LIX1      5.225919e-291  2.2180284 0.903 0.036 7.823723e-287
KCNN3     6.270610e-287  2.5730555 0.956 0.055 9.387730e-283
GPX3      3.004531e-282  3.1248437 0.929 0.050 4.498084e-278
MLC1      1.116737e-281  1.1081215 0.739 0.006 1.671866e-277
TMEM47    1.101784e-280  1.2469523 0.783 0.014 1.649481e-276
PLTP      9.259567e-272  1.7297954 0.867 0.034 1.386250e-267
AHNAK     1.655813e-267  1.5683290 0.872 0.034 2.478918e-263
S100A10   4.050418e-266  1.8971748 0.841 0.028 6.063881e-262
CD44      1.687907e-263  1.4189226 0.748 0.013 2.526965e-259
FAT3      7.289679e-263  1.1665198 0.735 0.013 1.091338e-258
SYTL4     7.397852e-259  1.6696984 0.916 0.058 1.107532e-254
TIMP3     1.566364e-253  1.9716725 0.898 0.050 2.345003e-249
CCDC80    2.290070e-248  1.0942316 0.681 0.009 3.428464e-244
PRLHR     7.525860e-247  1.3467037 0.765 0.025 1.126696e-242
CD99      1.292866e-246  3.0201008 0.996 0.109 1.935550e-242
NMB       1.921961e-245  4.0598647 0.951 0.091 2.877368e-241
BBOX1     3.804922e-244  0.8951846 0.637 0.004 5.696349e-240
EFEMP1    2.213013e-242  1.1109420 0.695 0.013 3.313102e-238




```{r, C11}
m.11 <- Seurat::FindMarkers(obj__diaz_2019__SF11136, ident.1 = 11)
head(m.11, n=25)
```


C1orf64  0.000000e+00  2.6756108 0.966 0.004  0.000000e+00
CDA      0.000000e+00  2.0107801 0.897 0.001  0.000000e+00
ADGRL4   0.000000e+00  1.8100046 0.828 0.001  0.000000e+00
CD34     0.000000e+00  1.5085390 0.862 0.001  0.000000e+00
ABCG2    0.000000e+00  3.0857655 1.000 0.005  0.000000e+00
ECSCR.1  0.000000e+00  1.3821448 0.793 0.001  0.000000e+00
DSP      0.000000e+00  0.8981521 0.759 0.000  0.000000e+00
ADGRF5   0.000000e+00  2.3632485 0.897 0.001  0.000000e+00
ABCB1    0.000000e+00  2.0047367 0.931 0.002  0.000000e+00
GIMAP7   0.000000e+00  1.2135070 0.862 0.001  0.000000e+00
GIMAP4   0.000000e+00  1.0532409 0.793 0.001  0.000000e+00
FLI1     0.000000e+00  1.1648225 0.862 0.002  0.000000e+00
ACVRL1   0.000000e+00  1.3625433 0.793 0.001  0.000000e+00
ATP10A   0.000000e+00  2.0513104 0.862 0.001  0.000000e+00
CGNL1    0.000000e+00  1.3298371 0.828 0.002  0.000000e+00
TMEM204  0.000000e+00  2.4958926 0.966 0.003  0.000000e+00
ICAM2    0.000000e+00  1.6439675 0.897 0.003  0.000000e+00
PCAT19   0.000000e+00  2.1880319 0.966 0.001  0.000000e+00
GGT5     0.000000e+00  2.4167906 1.000 0.005  0.000000e+00
ERG      0.000000e+00  1.5592281 0.862 0.000  0.000000e+00
PALMD   1.509625e-304  1.6576171 0.862 0.003 2.260059e-300
TM4SF1  5.269190e-304  2.3132205 0.931 0.004 7.888504e-300
SDPR    3.984029e-302  3.2381741 1.000 0.006 5.964490e-298
MECOM   5.141079e-297  1.2156861 0.793 0.002 7.696709e-293
TGM2    2.186670e-295  2.0348884 0.862 0.003 3.273664e-291



## Visualize



```{r}
Seurat::DimPlot(obj__diaz_2019__SF11136, reduction = "umap", label = TRUE, pt.size = .6, group.by = "celltype.annotated")
```



# Annotate componentents and/or clusters

```{r}


df <- -log10(geneset_component_associations(
  obj__diaz_2019__SF11136, 
  celltype_markers, 
  obj__diaz_2019__SF11136@misc$pca.informative.components
  ))

plt <- df |>
  tibble::rownames_to_column('PC')  |> 
  dplyr::mutate(order = as.numeric(gsub("^PC", "", PC))) |> 
  tidyr::pivot_longer(cols = -c('PC', 'order'), names_to = 'celltype') |> 
  dplyr::mutate(alpha = value > -log10(0.01))


ggplot(plt, aes(x=reorder(PC, order), y=value, fill=celltype, alpha=alpha)) +
  geom_bar(stat = "identity", position = "stack")

```


```{r}
ee <- data.frame()
for (pc in 1:34) {
  obj__diaz_2019__SF11136@reductions$pca@cell.embeddings[, pc:pc]

  t <- obj__diaz_2019__SF11136@reductions$pca@cell.embeddings[, pc:pc][obj__diaz_2019__SF11136$celltype == "tumor"]
  
  tam <- obj__diaz_2019__SF11136@reductions$pca@cell.embeddings[, pc:pc][obj__diaz_2019__SF11136$celltype == "TAM"]
  
  pe <- obj__diaz_2019__SF11136@reductions$pca@cell.embeddings[, pc:pc][obj__diaz_2019__SF11136$celltype == "pericyte"]
  
  en <- obj__diaz_2019__SF11136@reductions$pca@cell.embeddings[, pc:pc][obj__diaz_2019__SF11136$celltype == "endothelial"]

    ne <- obj__diaz_2019__SF11136@reductions$pca@cell.embeddings[, pc:pc][obj__diaz_2019__SF11136$celltype == "neuron"]
    
  tc <- obj__diaz_2019__SF11136@reductions$pca@cell.embeddings[, pc:pc][obj__diaz_2019__SF11136$celltype == "T-cell"]
  od <- obj__diaz_2019__SF11136@reductions$pca@cell.embeddings[, pc:pc][obj__diaz_2019__SF11136$celltype == "oligodendrocyte"]
  ac <- obj__diaz_2019__SF11136@reductions$pca@cell.embeddings[, pc:pc][obj__diaz_2019__SF11136$celltype == "astrocyte"]

  stats.t <- c(
    t.test(t, tam)$statistic,
    t.test(t, pe)$statistic,
    t.test(t, en)$statistic,
    t.test(t, ne)$statistic,
    t.test(t, tc)$statistic,
    t.test(t, od)$statistic,
    t.test(t, ac)$statistic
  )

  stats.tam <- c(
    t.test(tam, t)$statistic,
    t.test(tam, pe)$statistic,
    t.test(tam, en)$statistic,
    t.test(tam, ne)$statistic,
    t.test(tam, tc)$statistic,
    t.test(tam, od)$statistic,
    t.test(tam, ac)$statistic
  )

  stats.pe <- c(
    t.test(pe, t)$statistic,
    t.test(pe, tam)$statistic,
    t.test(pe, en)$statistic,
    t.test(pe, ne)$statistic,
    t.test(pe, tc)$statistic,
    t.test(pe, od)$statistic,
    t.test(pe, ac)$statistic
  )

  stats.en <- c(
    t.test(en, t)$statistic,
    t.test(en, tam)$statistic,
    t.test(en, pe)$statistic,
    t.test(en, ne)$statistic,
    t.test(en, tc)$statistic,
    t.test(en, od)$statistic,
    t.test(en, ac)$statistic
  )
  
    stats.tc <- c(
    t.test(tc, t)$statistic,
    t.test(tc, tam)$statistic,
    t.test(tc, pe)$statistic,
    t.test(tc, ne)$statistic,
    t.test(tc, en)$statistic,
    t.test(tc, od)$statistic,
    t.test(tc, ac)$statistic
  )
    
      
    stats.ne <- c(
    t.test(ne, t)$statistic,
    t.test(ne, tam)$statistic,
    t.test(ne, pe)$statistic,
    t.test(ne, tc)$statistic,
    t.test(ne, en)$statistic,
    t.test(ne, od)$statistic,
    t.test(ne, ac)$statistic
  )
    
          
    stats.od <- c(
    t.test(od, t)$statistic,
    t.test(od, tam)$statistic,
    t.test(od, pe)$statistic,
    t.test(od, tc)$statistic,
    t.test(od, en)$statistic,
    t.test(od, ne)$statistic,
    t.test(od, ac)$statistic
  )
          
    stats.ac <- c(
    t.test(ac, t)$statistic,
    t.test(ac, tam)$statistic,
    t.test(ac, pe)$statistic,
    t.test(ac, tc)$statistic,
    t.test(ac, en)$statistic,
    t.test(ac, ne)$statistic,
    t.test(ac, od)$statistic
  )


  ee <- rbind(
    ee,
    data.frame(
      pc = pc,
      assoc.t = sqrt(sum(stats.t^2)) ,
      assoc.tam = sqrt(sum(stats.tam^2)) ,
      assoc.pe = sqrt(sum(stats.pe^2)) ,
      assoc.en = sqrt(sum(stats.en^2)) ,
      assoc.tc = sqrt(sum(stats.tc^2)) ,
      assoc.ne = sqrt(sum(stats.ne^2)) ,
      assoc.od = sqrt(sum(stats.od^2)) ,
      assoc.ac = sqrt(sum(stats.ac^2)) 
    )
  )
}

ee

plt <- ee |> 
  tidyr::pivot_longer(-c('pc'))


ggplot(plt, aes(x=pc, y=value,fill=name)) + 
    geom_bar(stat = "identity", position = "fill")


```


## PC1
```{r}
Seurat::Idents(obj__diaz_2019__SF11136) <- "celltype.annotated"
Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = c("PCinv_1", "PC_1"),
  label = T,
  cols = c("red", "lightgrey", "blue")
)
Seurat::Idents(obj__diaz_2019__SF11136) <- "seurat_clusters"
```


```{r}
# top features
obj__diaz_2019__SF11136@reductions$pca@feature.loadings |> 
  as.data.frame() |> 
  dplyr::arrange(desc(abs(PC_1))) |> 
  dplyr::select("PC_1") |> 
  head(n=25)

FeaturePlot(obj__diaz_2019__SF11136, reduction="umap", features=c(""))

```

## PC2 - EN and/or PE
```{r}
Seurat::Idents(obj__diaz_2019__SF11136) <- "celltype.annotated"
Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = c("PCinv_2", "PC_2"),
  label = T,
  cols = c("red", "lightgrey", "blue")
)
Seurat::Idents(obj__diaz_2019__SF11136) <- "seurat_clusters"
```


## PC3 - tumor astrocyte like subcluster?
```{r}
Seurat::Idents(obj__diaz_2019__SF11136) <- "celltype.annotated"
Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = c("PCinv_3", "PC_3"),
  label = T,
  cols = c("red", "lightgrey", "blue")
)
Seurat::Idents(obj__diaz_2019__SF11136) <- "seurat_clusters"
```

```{r}
# top features
obj__diaz_2019__SF11136@reductions$pca@feature.loadings |> 
  as.data.frame() |> 
  dplyr::arrange(desc(abs(PC_3))) |> 
  dplyr::select("PC_3") |> 
  head(n=25)

Seurat::FeaturePlot(obj__diaz_2019__SF11136, reduction="umap", features=c("ID4"))
Seurat::FeaturePlot(obj__diaz_2019__SF11136, reduction="umap", features=c("GABRG1"))
```

## PC4 - TAM
```{r}
Seurat::Idents(obj__diaz_2019__SF11136) <- "celltype.annotated"
Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = c("PCinv_4", "PC_4"),
  label = T,
  cols = c("red", "lightgrey", "blue")
)
Seurat::Idents(obj__diaz_2019__SF11136) <- "seurat_clusters"
```


## PC5 - subcluster OD
```{r}
Seurat::Idents(obj__diaz_2019__SF11136) <- "celltype.annotated"
Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = c("PCinv_5", "PC_5"),
  label = T,
  cols = c("red", "lightgrey", "blue")
)
Seurat::Idents(obj__diaz_2019__SF11136) <- "seurat_clusters"
```


```{r}
# top features
obj__diaz_2019__SF11136@reductions$pca@feature.loadings |> 
  as.data.frame() |> 
  dplyr::arrange(desc(abs(PC_5))) |> 
  dplyr::select("PC_5") |> 
  head(n=25)

Seurat::FeaturePlot(obj__diaz_2019__SF11136, reduction="umap", features=c("TMEM108"))
Seurat::FeaturePlot(obj__diaz_2019__SF11136, reduction="umap", features=c("TMEM163"))
Seurat::FeaturePlot(obj__diaz_2019__SF11136, reduction="umap", features=c("PRICKLE1"))
```



## PC6 - EN vs PE ?
```{r}
Seurat::Idents(obj__diaz_2019__SF11136) <- "celltype.annotated"
Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = c("PCinv_6", "PC_6"),
  label = T,
  cols = c("red", "lightgrey", "blue")
)
Seurat::Idents(obj__diaz_2019__SF11136) <- "seurat_clusters"
```


```{r, top features}
obj__diaz_2019__SF11136@reductions$pca@feature.loadings |> 
  as.data.frame() |> 
  dplyr::arrange(desc(PC_6)) |> 
  dplyr::select("PC_6") |> 
  head(n=25)

obj__diaz_2019__SF11136@reductions$pca@feature.loadings |> 
  as.data.frame() |> 
  dplyr::arrange(PC_6) |> 
  dplyr::select("PC_6") |> 
  head(n=25)

Seurat::FeaturePlot(obj__diaz_2019__SF11136, reduction="umap", features=c("C1S"))
Seurat::FeaturePlot(obj__diaz_2019__SF11136, reduction="umap", features=c("DCN"))
Seurat::FeaturePlot(obj__diaz_2019__SF11136, reduction="umap", features=c("BTNL9"))
Seurat::FeaturePlot(obj__diaz_2019__SF11136, reduction="umap", features=c("CDA"))
```




# re-classification using markers
https://machinelearningmastery.com/linear-classification-in-r/

## LM https://topepo.github.io/caret/measuring-performance.html


## LDA


```{r}
obj__diaz_2019__SF11136 <- celltype_reclassification_lda(
  seurat_object = obj__diaz_2019__SF11136,
  celltype_markers |> dplyr::filter(celltype != "cell cycling") |> dplyr::pull(markers),
  "celltype.annotated"
)

```




```{r}
plot_sorted_classifier_scores(obj__diaz_2019__SF11136,'celltype.annotated' ,"^reclass.LDA.")
```



## LDA + PCA


```{r}
obj__diaz_2019__SF11136 <- celltype_reclassification_lda_with_pca(
  seurat_object = obj__diaz_2019__SF11136,
  celltype_markers |> dplyr::filter(celltype != "cell cycling") |> dplyr::pull(markers),
  20,
  "celltype.annotated"
)

```


```{r}
plot_sorted_classifier_scores(obj__diaz_2019__SF11136,'celltype.annotated' ,"^reclass.LDA_PCA.")
```


```{r}
Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = colnames(obj__diaz_2019__SF11136@meta.data)[grepl("^reclass.LDA_PCA.",colnames(obj__diaz_2019__SF11136@meta.data))],
  label = F,
  cols = c("lightgrey","red", "blue")
)
```


## PSLDA

```{r}
obj__diaz_2019__SF11136 <- celltype_reclassification_pslda(
  seurat_object = obj__diaz_2019__SF11136,
  celltype_markers |> dplyr::filter(celltype != "cell cycling") |> dplyr::pull(markers),
  20,
  "celltype.annotated"
)

obj__diaz_2019__SF11136 = flag_uncertain_classifications(obj__diaz_2019__SF11136, "PSLDA")
```

```{r}
#plot_sorted_classifier_scores(obj__diaz_2019__SF11136,'celltype.annotated' ,"^reclass\\.PSLDA\\.")
plot_sorted_classifier_scores(obj__diaz_2019__SF11136,'reclass.PSLDA.class' ,"^reclass\\.PSLDA\\.", "noskip$")
```


```{r}
Seurat::FeaturePlot(obj__diaz_2019__SF11136,
  reduction = "umap",
  features = colnames(obj__diaz_2019__SF11136@meta.data)[grepl("^reclass\\.PSLDA\\.",colnames(obj__diaz_2019__SF11136@meta.data))],
  label = F,
  cols = c("lightgrey","red", "blue")
)
```

### min classifier scores

#### PE

```{r}
plot_sorted_classifier_scores(obj__diaz_2019__SF11136,'celltype.annotated' ,"^reclass\\.PSLDA\\.") + 
  geom_hline(yintercept = obj__diaz_2019__SF11136@misc$reclass.PSLDA.min.scores$`pericyte`, lwd=0.5, col="darkgray")
```


#### OD


```{r}
plot_sorted_classifier_scores(obj__diaz_2019__SF11136,'celltype.annotated' ,"^reclass\\.PSLDA\\.") + 
  geom_hline(yintercept = obj__diaz_2019__SF11136@misc$reclass.PSLDA.min.scores$`oligodendrocyte`, lwd=0.5, col="darkgray")
```



#### T ~= 0.33



```{r}
plot_sorted_classifier_scores(obj__diaz_2019__SF11136,'celltype.annotated' ,"^reclass\\.PSLDA\\.") + 
  geom_hline(yintercept = obj__diaz_2019__SF11136@misc$reclass.PSLDA.min.scores$`tumor`, lwd=0.5, col="darkgray")
```



## SVM 


## SVM + PCA





# Store object

## Check git code changes and save

```{r}
cache_seurat_object(obj__diaz_2019__SF11136)
```




# Visualizations

```{r}
FeaturePlot(obj__diaz_2019__SF11136, features=c("TMEM144"))

scripts/scRNA_analyses_Yuan.R:FeaturePlot(object = object_1, features = "MOG")
scripts/scRNA_analyses_Yuan.R:FeaturePlot(object = object_1, features = "PLP1")
scripts/scRNA_analyses_Yuan.R:FeaturePlot(object = object_1, features = "TMEM144")
scripts/scRNA_analyses_Yuan.R:FeaturePlot(object = object_1, features = "TMEM125")
scripts/scRNA_analyses_Yuan.R:FeaturePlot(object = object_1, features = "ABCB1")
scripts/scRNA_analyses_Yuan.R:FeaturePlot(object = object_1, features = "CD34")
scripts/scRNA_analyses_Yuan.R:FeaturePlot(object = object_1, features = "FLT4")
scripts/scRNA_analyses_Yuan.R:FeaturePlot(object = object_1, features = "TIE1") # meh
scripts/scRNA_analyses_Yuan.R:FeaturePlot(object = object_1, features = "ITGA1") # endo + peri?
scripts/scRNA_analyses_Yuan.R:FeaturePlot(object = object_1, features = "RGS5")
scripts/scRNA_analyses_Yuan.R:FeaturePlot(object = object_1, features = "PDGFRB")
scripts/scRNA_analyses_Yuan.R:FeaturePlot(object = object_1, features = "CD248")


```


